---
title: "Common garden so far"
author: "Jane"
date: "11/27/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Reading in and cleaning up the data.

```{r }
library(ggplot2)
library(googlesheets4)
library(tidyverse)
#Register sheet
G <- read_sheet(ss="https://docs.google.com/spreadsheets/d/1ZbeKlGR9pUXoUQO2xaGx0Jt_Z1c5zNjrI_1Cgi3ZuDI/edit#gid=550678747", sheet="CG_measurements")

#Filter for only seedlings included in the study
G <- G[G$included=="y",]
G$type <- as.factor(G$type)
G$mom <-as.factor(G$mom)

#Dataframe of native seedlings
n <- G[G$type=="n",]
#Dataframe of cultivar seedlings
c <- G[G$type=="c",]

#All columns with equations in them will read in as a list rather than numeric. Haven't figured out how to avoid that yet, so I'm just creating the column I need here.
G$Height_cm9.18.19 <- G$Height_mm9.18.19/10

```

The variables:
stem_base_mm5.30.18 : collected 5/30/18 shortly after the seedlings were planted in the common garden
stem_below_contyledons_mm5.30.18 : collected 5/30/18, yes there's a typo in the name
stemd2.18.19 : collected 2/18/19, clearly after the first growing season
Height_cm6.24.18: height near the beginning of the first growing season
Height_cm10.20.18: collected 10/20/19 at the end of the first growing season

Summary statistics

```{r}
#Means
#Height
h62418c <-mean(c$Height_cm6.24.18, na.rm=TRUE)#beginning of 2018
h62418n <-mean(n$Height_cm6.24.18, na.rm=TRUE)
h102018c <-mean(c$Height_cm10.20.18, na.rm=TRUE)#end of 2018
h102018n <-mean(n$Height_cm10.20.18, na.rm=TRUE)
h5219c <-mean(c$Height_cm5.2.19, na.rm=TRUE)#beginning of 2019
h5219n <-mean(n$Height_cm5.2.19, na.rm=TRUE)
h91819c <-mean((c$Height_mm9.18.19/10), na.rm=TRUE)#end of 2019
h91819n <-mean((n$Height_mm9.18.19/10), na.rm=TRUE)

#differences in the mean height over time
height <-data.frame(type=c(rep("native",4),rep("cultivar",4)),height=c(h62418n,h102018n,h5219n,h91819n,h62418c,h102018c,h5219c,h91819c), census=c(1:4))

h <- ggplot(height) + geom_line(aes(x=census, y=height, color=type))+theme_bw()+ylab("Height")
h


#Stem diameter
sd53018c <- mean(c$stem_below_cotyledons_mm5.30.18, na.rm=TRUE)#beginning of 2018
sd53018n <- mean(n$stem_below_cotyledons_mm5.30.18, na.rm=TRUE)
sd21819c <- mean(c$stemd2.18.19, na.rm=TRUE)#End of season 1
sd21819n <- mean(n$stemd2.18.19, na.rm=TRUE)
sd5219c <- mean(c$stemd5.2.19, na.rm=TRUE)#Beginning of season 2
sd5219n <- mean(n$stemd5.2.19, na.rm=TRUE)
sd91719c <- mean(c$stemd9.17.19, na.rm=TRUE)#End of season 2
sd91719n <- mean(n$stemd9.17.19, na.rm=TRUE)

```

##**Stem diameter**

```{r}
library(nlme)
library(multcomp)
library(lmerTest)

#Beginning of the experiment
#Stemdiameter 5/30/18
#Get rid of those NA values
g <-G %>% drop_na(stem_base_mm5.30.18)
g$stem_base_mm5.30.18 <-scale(g$stem_base_mm5.30.18, center=TRUE,scale = TRUE)
stemres53018 <-lm(g$stem_base_mm5.30.18~g$days)
g$stemres53018<-stemres53018$residuals
#residuals
qqnorm(g$stemres53018)
shapiro.test(g$stemres53018)

#Model with May 30th, 2018 stem diameter residuals as afunction of mom and type
mod1 <- lmer(g$stemres53018~(1|g$mom)+g$type)
summary(mod1)
#Model with stem residuals as a function of mom and type but a different function
mod2 <- lme(stemres53018~type,data=g, random= ~ type|mom)
summary(mod2)
#Model withs stem diameter as a funciton of grow days, mom, and type
mod <- lmer(g$stem_base_mm5.30.18~g$days+(1|g$mom)+g$type)
summary(mod)

#End of growning season 1
#Stemdiameter 2/18/19
#Get rid of those NA values
g <-G %>% drop_na(stemd2.18.19)
#Regress out the grow days
stemres21819 <-lm(g$stemd2.18.19~g$days)
#Add the residuals to the data frame
g$stemres21819<-stemres21819$residuals
#Looking at theresiduals
qqnorm(g$stemres21819)
shapiro.test(g$stemres21819)
#Model with Feb 18, 2019 stem residuals as a function of mom and type
mod1 <- lmer(g$stemres21819~(1|g$mom)+g$type)
summary(mod1)
#Model with Feb 18, 2019 stem residuals as a function of mom and type with a different function
mod2 <- lme(stemres21819~type,data=g, random= ~ type|mom)
summary(mod2)
#Model with Feb 18, 2019 stem diameter as a function of mom and type
mod <- lmer(g$stemd2.18.19~g$days+(1|g$mom)+g$type)
summary(mod)


#Near the beginning of growing season 2
#Stemdiameter 5/2/19
#Get rid of those NA values
g <-G %>% drop_na(stemd5.2.19)

stemres5219 <-lm(g$stemd5.2.19~g$days)
g$stemres5219<-stemres5219$residuals

qqnorm(g$stemres5219)
shapiro.test(g$stemres5219)

mod1 <- lmer(g$stemres5219~(1|g$mom)+g$type)
summary(mod1)

mod <- lmer(g$stemd5.2.19~g$days+(1|g$mom)+g$type)
summary(mod)

#mod2 <- lme(stemres5219~type,data=g, random= ~ type|mom)
#summary(mod2)

#Get rid of those NA values
#Stemdiameter 9/17/19
g <-G %>% drop_na(stemd9.17.19)

stemres91719 <-lm(g$stemd9.17.19~g$days)
g$stemres<-stemres91719$residuals

qqnorm(g$stemres)
shapiro.test(g$stemres)

mod1 <- lmer(g$stemres~(1|g$mom)+g$type)
summary(mod1)

mod <- lmer(g$stemd9.17.19~g$days+(1|g$mom)+g$type)
summary(mod)

#mod2 <- lme(stemres~type,data=g, random= ~ type|mom)
#summary(mod2)
```

###Non-nested, non-parametric test.

```{r}
#non-nested
anov <-aov(stemres~type, g)
summary(anov)

#non-parametric
kruskal.test(stemres~type,g)

```


##**Height**

```{r}
#Height 6/24/18
#Get rid of those NA values
g <-G %>% drop_na(Height_cm6.24.18)

h62418res <-lm(g$Height_cm6.24.18~g$days)
g$h62418res<-h62418res$residuals

qqnorm(g$h62418res)
shapiro.test(g$h62418res)

mod1 <- lmer(g$h62418res~(1|g$mom)+g$type)
summary(mod1)

mod <- lmer(g$Height_cm6.24.18~g$days+(1|g$mom)+g$type)
summary(mod)

mod2 <- lme(h62418res~type,data=g, random= ~ type|mom)
summary(mod2)

#Height 10/20/18
g <-G %>% drop_na(Height_cm10.20.18)

h102018res <-lm(g$Height_cm10.20.18~g$days)
g$h102018res<-h102018res$residuals

qqnorm(g$h102018res)
shapiro.test(g$h102018res)

mod1 <- lmer(g$h102018res~(1|g$mom)+g$type)
summary(mod1)

mod <- lmer(g$Height_cm10.20.18~g$days+(1|g$mom)+g$type)
summary(mod)

#mod2 <- lme(h102018res~type,data=g, random= ~ type|mom)
#summary(mod2)

#Height 5/2/19
g <-G %>% drop_na(Height_cm5.2.19)

h5219res <-lm(g$Height_cm5.2.19~g$days)
g$h5219res<-h5219res$residuals

qqnorm(g$h5219res)
shapiro.test(g$h5219res)

mod1 <- lmer(g$h5219res~(1|g$mom)+g$type)
summary(mod1)

mod <- lmer(g$Height_cm5.2.19~g$days+(1|g$mom)+g$type)
summary(mod)

mod2 <- lme(h5219res~type,data=g, random= ~ type|mom)
summary(mod2)


#Height 9/18/19
g <-G %>% drop_na(Height_cm9.18.19)

g$daysln <- log(g$days)
h91819res <-lm(g$Height_cm9.18.19~g$days)
g$h91819res<-h91819res$residuals

qqnorm(g$h91819res)
shapiro.test(g$h91819res)

mod1 <- lmer(g$h91819res~(1|g$mom)+g$type)
summary(mod1)

mod <- lmer(g$Height_cm9.18.19~g$days+(1|g$mom)+g$type)
summary(mod)

mod2 <- lme(h91819res~type,data=g, random= ~ type|mom)
summary(mod2)
```


##**Herbivory total

```{r}
g <-G %>% drop_na(Herbivory7.22.19)
#Herbivory as a function of mom and type
mod1 <- lmer(g$Herbivory7.22.19~(1|g$mom)+g$type)
summary(mod1)

herb <-ggplot(g)
herb <- herb + geom_boxplot(na.rm=TRUE,aes(x=factor(type),y=Herbivory7.22.19, color=type), notch = FALSE)+theme_bw()+ylab("% herbivory")+ylim(0,100)
herb <- herb +scale_color_manual(values=c("#006622", "#00b33c"))+ theme(panel.background = element_rect(fill = "white"),plot.background = element_rect(fill = "transparent",colour = NA),text = element_text(size=20))
herb <- herb +theme(legend.position = "none", axis.title.x=element_blank())
herb <- herb + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
herb
#ggsave(herb, filename = "herbivory.png")

```

##**Number of leaves

```{r}
#Number of leaves beginning of season
g <-G %>% drop_na(leafcount7.9.19)
mod <- lmer(g$leafcount7.9.19~(1|g$mom)+g$type)
summary(mod)

#g$leafcount7.9.19sq <-sqrt(g$leafcount7.9.19)

boxl <-ggplot(g)
boxl <- boxl + geom_boxplot(na.rm=TRUE,aes(x=factor(type),y=leafcount7.9.19, color=type), notch = TRUE)+theme_bw()+ylab("Number of leaves")+ylim(0,40)
boxl <- boxl +scale_color_manual(values=c("#006622", "#00b33c"))+ theme(panel.background = element_rect(fill = "white"),plot.background = element_rect(fill = "transparent",colour = NA),text = element_text(size=20))
boxl <- boxl +theme(legend.position = "none", axis.title.x=element_blank())
boxl <- boxl + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
boxl
#ggsave(boxl, filename = "leavesJuly.png")

#Number of leaves end of season
g <-G %>% drop_na(Total_leaves09.08.19)

mod <- lmer(g$Total_leaves09.08.19~(1|g$mom)+g$type)
summary(mod)

#g$Total_leaves09.08.19sq <- sqrt(g$Total_leaves09.08.19)

boxl1 <-ggplot(g)
boxl1 <- boxl1 + geom_boxplot(na.rm=TRUE,aes(x=factor(type),y=Total_leaves09.08.19, color=type), notch = TRUE)+theme_bw()+ylab("Number of leaves")+ylim(0,40)
boxl1 <- boxl1 +scale_color_manual(values=c("#006622", "#00b33c"))+ theme(panel.background = element_rect(fill = "white"),plot.background = element_rect(fill = "transparent",colour = NA),text = element_text(size=20))
boxl1 <- boxl1 +theme(legend.position = "none", axis.title.x=element_blank())
boxl1 <- boxl1 + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
boxl1
#ggsave(boxl1, filename = "leavesSept.png")

```

##**Number of Secondary Branches
```{r}
#Number of leaves end of season
g <-G %>% drop_na(num_secondarybranches)

mod <- lmer(g$num_secondarybranches~(1|g$mom)+g$type)
summary(mod)

#g$num_secondarybranchessq <- sqrt(g$num_secondarybranches)

boxl1 <-ggplot(g)
boxl1 <- boxl1 + geom_boxplot(na.rm=TRUE,aes(x=factor(type),y=num_secondarybranches, color=type), notch = FALSE)+theme_bw()+ylab("Number of secondary branches")+ylim(0,5)
boxl1 <- boxl1 +scale_color_manual(values=c("#006622", "#00b33c"))+ theme(panel.background = element_rect(fill = "white"),plot.background = element_rect(fill = "transparent",colour = NA),text = element_text(size=20))
boxl1 <- boxl1 +theme(legend.position = "none", axis.title.x=element_blank())
boxl1 <- boxl1 + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
boxl1
#ggsave(boxl1, filename = "Secondarybranches.png")


```

## Plots

Plots

```{r pressure, echo=FALSE}

#Boxplot Stem diameter 5/30/18
boxs1 <-ggplot(G)
boxs1 <- boxs1 + geom_boxplot(aes(x=factor(type),y=stem_base_mm5.30.18, color=type), notch = TRUE)+theme_bw()+ylab("Stem diameter (mm)")+ylim(0,4)
boxs1 <- boxs1 +scale_color_manual(values=c("#006622", "#00b33c"))+ theme(panel.background = element_rect(fill = "white"),plot.background = element_rect(fill = "transparent",colour = NA),text = element_text(size=20))
boxs1 <- boxs1 +theme(legend.position = "none", axis.title.x=element_blank())
boxs1 <- boxs1 + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
boxs1
#ggsave(boxs1, filename = "stem_base53018.png")

#Stem diameter 2/18/19
boxs2 <-ggplot(G)
boxs2 <- boxs2 + geom_boxplot(na.rm=TRUE,aes(x=factor(type),y=stemd2.18.19, color=type), notch = TRUE)+theme_bw()+ylab("Stem diameter (mm)")+ylim(0,4)
boxs2 <- boxs2 +scale_color_manual(values=c("#006622", "#00b33c"))+ theme(panel.background = element_rect(fill = "white"),plot.background = element_rect(fill = "transparent",colour = NA),text = element_text(size=20))
boxs2 <- boxs2 + theme(legend.position = "none", axis.title.x=element_blank())
boxs2 <- boxs2 + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
boxs2
#ggsave(boxs2, filename = "stem_diameter21819.png")

#Stem diameter 5/2/19
boxs3 <-ggplot(G)
boxs3 <- boxs3 + geom_boxplot(na.rm=TRUE,aes(x=factor(type),y=stemd5.2.19, color=type), notch = TRUE)+theme_bw()+ylab("Stem diameter (mm)")+ylim(0,4)
boxs3 <- boxs3 +scale_color_manual(values=c("#006622", "#00b33c"))+ theme(panel.background = element_rect(fill = "white"),plot.background = element_rect(fill = "transparent",colour = NA),text = element_text(size=20))
boxs3 <- boxs3 + theme(legend.position = "none", axis.title.x=element_blank())
boxs3 <- boxs3 + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
boxs3
#ggsave(boxs3, filename = "stem_diameter5219.png")

#Stem diameter 9/17/19
boxs4 <-ggplot(G)
boxs4 <- boxs4 + geom_boxplot(na.rm=TRUE,aes(x=factor(type),y=stem_diameter091719, color=type), notch = TRUE)+theme_bw()+ylab("Stem diameter (mm)")+ylim(0,4)
boxs4 <- boxs4 +scale_color_manual(values=c("#006622", "#00b33c"))+ theme(panel.background = element_rect(fill = "white"),plot.background = element_rect(fill = "transparent",colour = NA),text = element_text(size=20))
boxs4 <- boxs4 + theme(legend.position = "none", axis.title.x=element_blank())
boxs4 <- boxs4 + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
boxs4
#ggsave(boxs4, filename = "stem_diameter91719.png")


#Height 6/24/18
boxh <-ggplot(G)
boxh <- boxh + geom_boxplot(na.rm=TRUE,aes(x=factor(type),y=Height_cm6.24.18, color=type), notch = TRUE)+theme_bw()+ylab("Plant Height (cm)")+ylim(0,30)
boxh <- boxh +scale_color_manual(values=c("#006622", "#00b33c"))+ theme(panel.background = element_rect(fill = "white"),plot.background = element_rect(fill = "transparent",colour = NA),text = element_text(size=20))
boxh <- boxh + theme(legend.position = "none", axis.title.x=element_blank())
boxh <- boxh + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
boxh
#ggsave(boxh, filename = "height62418.png")

#Height 9/18/19
boxh1 <-ggplot(G)
boxh1 <- boxh1 + geom_boxplot(aes(x=factor(type),y=Height_cm9.18.19, color=type), notch = TRUE)+theme_bw()+ylab("Plant Height (cm)")+ylim(0,30)
boxh1 <- boxh1 +scale_color_manual(values=c("#006622", "#00b33c"))+ theme(panel.background = element_rect(fill = "white"),plot.background = element_rect(fill = "transparent",colour = NA),text = element_text(size=20))
boxh1 <- boxh1 + theme(legend.position = "none", axis.title.x=element_blank())
boxh1 <- boxh1 + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
boxh1
#ggsave(boxh1, filename = "height91819.png")



```


#Survival

```{r}
first <-G %>% drop_na(stem_base_mm5.30.18) #plants alive 5/30/18
fn <- nrow(first[first$type=="n",])
fc <- nrow(first[first$type=="c",])
second <- G %>% drop_na(Height_cm6.24.18) #plants alive 6/24/18
sn <- nrow(second[second$type=="n",])
sc <- nrow(second[second$type=="c",])
third <-G %>% drop_na(Height_cm10.20.18) #plants alive 10/20/18
tn <- nrow(third[third$type=="n",])
tc <- nrow(third[third$type=="c",])
fourth <-G %>% drop_na(Height_cm5.2.19) #plants alive 5/2/19
Fn <- nrow(fourth[fourth$type=="n",])
Fc <- nrow(fourth[fourth$type=="c",])
fifth <-G %>% drop_na(Height_cm9.18.19) #plants alive 9/18/19
ffn <- nrow(fifth[fifth$type=="n",])
ffc <- nrow(fifth[fifth$type=="c",])


#Percent survival
n1 <-fn/fn*100
c1 <-fc/fc*100
n2 <-sn/fn*100
c2 <-sc/fc*100
n3 <-tn/fn*100
c3 <-tc/fc*100
n4 <-Fn/fn*100
c4 <-Fc/fc*100
n5 <-ffn/fn*100
c5 <-ffc/fc*100


native <- data.frame(ncensus=c(1:5), nsurvival=c(n1,n2,n3,n4,n5))
cultivar <- data.frame(ccensus=c(1:5), csurvival=c(c1,c2,c3,c4,c5))



survive <- glmer(survival~type+(1|mom), data=first, family="binomial")
summary(survive)

nrow(fifth[fifth$type=="n",])/nrow(n)*100
nrow(fifth[fifth$type=="c",])/nrow(c)*100

```

#Survival plots

```{r}
census <-data.frame(num=c(1:5),date=c(1:5),survival=c(n1,n2,n3,n4,n5,c1,c2,c3,c4,c5),type=c(rep("native",5),rep("cultivar",5)))


g <- ggplot(census) + geom_line(aes(x=date, y=survival, color=type))+theme_bw()+ylab("% Survival")
g <-g+ theme(panel.background = element_rect(fill = "white"),plot.background = element_rect(fill = "transparent",colour = NA),text = element_text(size=20))
g <-g + theme(legend.justification=c(1,1), legend.position=c(0.25,0.35)) + ylim(75,100)
g <- g +theme(axis.title.x=element_blank())+scale_color_manual(values=c("#006622", "#00b33c"))
g <- g + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
g

#ggsave(g, filename = "survival.png")
```


##**Biomass**##

```{r}
#Register sheet
a <- read_sheet(ss="https://docs.google.com/spreadsheets/d/1zQ7bHuapb72X1cdpxU3ix6otKdnox51Numy6cUaS97g/edit#gid=0", sheet="CG_biomass")

#Filter for only seedlings included in the study
a <- a[a$included=="y",]
a$type <- as.factor(a$type)
a$mom <-as.factor(a$mom)
a$stemt <-a$stem_AGB+a$Stem
#Dataframe of native seedlings
n <- a[a$type=="n",]
#Dataframe of cultivar seedlings
c <- a[a$type=="c",]

AGBlc <-mean(c$leaf_AGB,na.rm=TRUE)
AGBln <-mean(n$leaf_AGB,na.rm=TRUE)
AGBsc <-mean(c$stemt,na.rm=TRUE)
AGBsn <-mean(n$stemt,na.rm=TRUE)

#Leaf AGB
A <-a %>% drop_na(leaf_AGB)

mod <- lmer(A$leaf_AGB~(1|A$mom)+A$type)
summary(mod)

abg <-ggplot(a)
abg <- abg + geom_boxplot(na.rm=TRUE,aes(x=factor(type),y=leaf_AGB, color=type), notch = TRUE)+theme_bw()+ylab("Leaf AGB (g)")+ylim(0,3)
abg <- abg +scale_color_manual(values=c("#006622", "#00b33c"))+ theme(panel.background = element_rect(fill = "white"),plot.background = element_rect(fill = "transparent",colour = NA),text = element_text(size=20))
abg <- abg +theme(legend.position = "none", axis.title.x=element_blank())
abg <- abg + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
abg
#ggsave(abg, filename = "leafabg.png")

#Stem AGB
A <-a %>% drop_na(stem_AGB)

mod <- lmer(A$stem_AGB~(1|A$mom)+A$type)
summary(mod)

abg <-ggplot(a)
abg <- abg + geom_boxplot(na.rm=TRUE,aes(x=factor(type),y=stem_AGB, color=type), notch = TRUE)+theme_bw()+ylab("Stem AGB (g)")+ylim(0,3)
abg <- abg +scale_color_manual(values=c("#006622", "#00b33c"))+ theme(panel.background = element_rect(fill = "white"),plot.background = element_rect(fill = "transparent",colour = NA),text = element_text(size=20))
abg <- abg +theme(legend.position = "none", axis.title.x=element_blank())
abg <- abg + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
abg
#ggsave(abg, filename = "stemabg.png")

```
